Loss recovery and tail loss performance in Multipath TCP

### 2016-12-23 main.tex
edit

\section{Introduction}
TCP has two mechanisms for detecting and recovering from packet losses namely Fast Recovery (FR) and Retransmission Timeouts(RTOs). For long flows where there are sufficiently large number of packets 
to be transferred, FR is quicker than RTOs in detecting and recovering packet losses. On the other hand shorter flows which are a majority in the web traffic have to depend on RTOs for packet loss detection 
and recovery incurring more latency. A single packet loss in a short flow may take many RTTs to detect and recover. This scenario is also applicable to the packets at the end of the flow (or tail) in a long 
flow.

@@

\begin{table}
\begin{center}
\begin{tabular}{|c|cccccccccc|}
      \hline
      \multicolumn{1}{c}{} & & \\[\dimexpr-\normalbaselineskip-\arrayrulewidth]
      \textbf{Burst Size} & \multicolumn{10}{c|}{160 Packets} \\
      \hline
      \textbf{Separation Time} & \multicolumn{10}{c|}{2s} \\
      \hline

@@

\subsection{Testing with Deterministic loss patterns}
In order to understand the retransmission behavior of the Linux MPTCP implementation and to reproduce the observed retransmissions, we use a deterministic drop pattern.
Losses generated by using associating netem with corresponding interfaces and dropping the packets. This process is simplified by using the KAUNetem tool~\cite{Garcia2016}. 
MPTCP connection starts with a single TCP subflow and subsequently, one or more subflows are added following an agreement between client and server on the MPTCP protocol 
support and interface availability. So one has to wait in time and packets for the second subflow establishment to understand the MPTCP retransmissions. We send two bursts of 80 
packets each and drop tail packets on one of the interface. This is to ensure that the necessary and sufficient conditions for probe triggering are met and tail loss probe is generated on that
subflow. Three test cases are evaluated with one packet tail loss, two packet tail loss and  one packet along with probe loss.

In Linux, TCP retransmission features are controlled by a sysctl setting tcp.early.retrans. It has 5 possible values with ranging from 0 to 4 with 3 being Linux default, enables both
ER and TLP. For this analysis, we used 0,3 and 4. In each setting, we calculate burst completion time for two burst of 80 packets each. The setup has server and client with MPTCP
enabled linux running on them. Client has two wireless interfaces with one way delay on each interface ranging from 20ms to 120ms. We consider 5 scenarios with delay pairings 20ms-20ms
20ms-30ms, 20ms-120ms, 30ms-20ms, 120ms-20ms to understand the effect of delay difference in the performance.

Figures ~\ref{t1p,t2p,t1pp} represent the comparison of TCP burst completion times for the two bursts from server to client.

@@

The expected retransmission behavior for different settings of ER is shown 
in~\ref{timingER0},~/\ref{timingER3} and ~\ref{timingER4}.

@@

@@

\begin{figure}[!ht]
\begin{center}
\includegraphics[angle=0, width=0.45\textwidth, natwidth=610, natheight=400]{images/timingER4.pdf}
\end{center}
\caption{Timing diagram of different tail loss scenarios in sysctl ER=4}\label{timingER4}
\end{figure}

>>>>

\section{Introduction}
TCP has two mechanisms for detecting and recovering from packet losses namely Fast Recovery (FR) and Retransmission Timeouts(RTOs). For long flows where there are sufficiently large number of packets 
to be transferred, FR is quicker than RTOs in detecting and recovering packet losses. On the other hand shorter flows which are a majority in the web traffic have to depend on RTOs for packet loss detection 
and recovery thus incurring more latency. A single packet loss in a short flow may take many RTTs to detect and recover. This scenario is also applicable to the packets at the end of the flow (or tail) in a long 
flow.

@@

\begin{table}
\begin{center}
\begin{tabular}{|c|cccccccccc|}
      \hline
      \multicolumn{1}{c}{} & & \\[\dimexpr-\normalbaselineskip-\arrayrulewidth]
      \textbf{Burst Size} & \multicolumn{10}{c|}{80 Packets} \\
      \hline
      \textbf{Separation Time} & \multicolumn{10}{c|}{2s} \\
      \hline

@@

\subsection{Testing retransmission with deterministic loss patterns}
In order to understand the retransmission behavior of the Linux MPTCP implementation and to reproduce the observed retransmissions, we use a deterministic drop pattern.
Losses generated by using associating netem with corresponding interfaces and dropping the packets. This process is simplified by using the KAUNetem tool~\cite{Garcia2016}. 
MPTCP connection starts with a single TCP subflow and subsequently, one or more subflows are added following an agreement between client and server on the MPTCP protocol 
support and interface availability. So one has to wait in time and packets for the second subflow establishment to understand the MPTCP retransmissions. We send two bursts of 80 
packets each and drop tail packets on one of the interface. This is to ensure that the necessary and sufficient conditions for probe triggering are met and tail loss probe is generated on that
subflow. Three test cases are evaluated with one packet tail loss, two packet tail loss and  one packet along with probe loss.

In Linux, TCP retransmission features are controlled by a sysctl setting tcp.early.retrans. It has 5 possible values with ranging from 0 to 4 with 3 being Linux default, enables both
ER and TLP. For this analysis, apart from the default setting, we also consider 0 that diasables both ER and TLP and 2 that enables delayed ER. Disabling ER and TLP gives us the performance
of RTO efficiency. Enabling delayed ER gives us combined performance of RTO and delayed ER and provides the possibility to single out the scenarios where TLP is more effective.
Probe loss should be seen as retransmission loss when tcp.early.retrans is 0 or 2.

In each setting, we calculate burst completion time for two bursts of 80 packets each. The setup has server and client with Linux supporting MPTCP running on them. Client has two wireless 
interfaces with one way delay on each interface ranging from 20ms to 120ms. We consider 5 scenarios with delay pairings 20ms-20ms, 20ms-30ms, 20ms-120ms, 30ms-20ms, 120ms-20ms to understand 
the effect of delay difference in the performance. These scenarios are choosen to evaluate the effect of symmetry, mild asymmetry and higher asymmetry in link delays.

Retransmission behavior of TCP is seen a base case for MPTCP performance as each individual MPTCP subflow is a TCP flow in itself. Our analysis
starts with the results using TCP and compares it with that of the MPTCP. Figures ~\ref{t1p},~\ref{t2p},~\ref{t1pp} represent the comparison of TCP 
burst completion times.

@@

The expected retransmission behavior of MPTCP for different settings of ER is shown 
in~\ref{timingER0},~\ref{timingER2} and ~\ref{timingER3}.

@@

\begin{figure}[!ht]
\begin{center}
\includegraphics[angle=0, width=0.45\textwidth, natwidth=610, natheight=400]{images/timingER2.pdf}
\end{center}
\caption{Timing diagram of different tail loss scenarios in sysctl ER=2}\label{timingER2}
\end{figure}

@@
